Parameter Space Representation of Pareto Front to Explore Hardware-Software Dependencies

### 2014-07-28 introduction.tex
readded pareto descr

The design of an embedded system requires different conflicting
objectives (energy consumption, performance, area) to be fulfilled.
Several factors are involved in this goal and real cases show how
often no analytical results are known to predict the relation between
system configurations and objectives. In such cases a high level
system simulation is the only way to have a picture of how system
parameters impact on the objectives.  A key problem is that the size
of the design space to be simulated grows with the product of
cardinalities of each parameter, resulting in an intractable number
of simulations. In these cases a Design Space
Exploration (DSE) strategy is required, i.e. an algorithm that selects
only a tractable subset of all possible configurations to simulate while providing, at the end, configurations that are close to the optimum and that are worth evaluating, in a following step of the design process, in a more thorough way.

%The final result of a design space exploration is a subset of the configurations called \emph{Pareto set} (\cite{pareto}, see also \defR{pareto-set}). The corresponding values of the objectives for each configuration of the Pareto Set is called \emph{Pareto front}: it represents the trade-offs between objectives, i.e.  for each point of the Pareto front there is no other configuration performing better for all the objectives considered. Once these few promising configurations have been obtained, a subsequent step of the design flow can eventually afford a more accurate low level simulation.

In this work we present a multiobjective design exploration algorithm
that introduces the concept of interesting and uninteresting regions of the configuration space.
\comment{AA: prima era ``algorithm
that introduces the concept of Parameter Space Representation of
Pareto front''. Se io fossi un reviewer, direi: guardi, questo io non credo, anche pecche' non siamo noi che introduciamo il concetto di parameter space representation del pareto set. Il concetto esiste gia' dal 1896 e si chiama ``Pareto-set''! Ho cambiato il claim dicendo che introduciamo interesting/uninteresting regions, che mi sembra piu' rispondente alla verita'. Cosi' ti va bene?} 
The algorithm is iterative. The main idea can be simplified as follows. At each iteration, the configuration space is divided in regions. A score is given to each region based on the \emph{innovation} provided by its configurations in the Pareto front w.r.t. the previous iterations. The regions with high score are considered the most \emph{interesting} and more simulations are distributed there rather than in the rest of the configuration space. Since the fundamental operations consist in the formation of regions in the Parameter Space (aka configuration space), we call our algorithm PS.

\comment{AA: Le seguenti referenze a cose di psicologia cognitiva potrebbero essere interpretate come una figata o una cagata. Tu come le interpreti?}
Our algorithm mimics the attention shift in human beings. Authors in \cite{attention} models human attention as a single resource with a limited budget that must be distributed among tasks, preferring the task which the focus is on.
 Authors in \cite{spatial_attention} describes the visual attention as a spotlight that focuses, each time, in a particular region of the visual space in order to process it more deeply. Similarly, PS focuses its ``spotlight'' only to the interesting regions and explore them more thoroughly distributing the simulation budget on them.
% Ispirzione: http://en.wikipedia.org/wiki/Attentional_shift

With PS, we propose a new
perspective on the configuration space, where parameter
interdependencies play the fundamental role of making some regions
more capable of generating hardly predictable solutions: catching this
sort of ``chaoticity'' in the parameter space is the main idea behind
the strategy that is being introduced in this work.

>>>>

The design of an embedded system requires different conflicting
objectives (energy consumption, performance, area) to be fulfilled.
The growing demand and reduced time-to-market of these systems are
introducing even more challenges that designers have to
face~\cite{wsts}.
Several factors are involved in this goal and real cases show how
often no analytical results are known to predict the relation between
system configurations and objectives. In such cases a high level
system simulation is the only way to have a picture of how system
parameters impact on the objectives.  A key problem is that the size
of the design space to be simulated grows with the product of
cardinalities of each parameter, resulting in an intractable number
of simulations.

In these cases a Design Space
Exploration (DSE) strategy is required, i.e. an algorithm that selects
only a tractable subset of all possible configurations.
The final result of a design space exploration is a subset of the
configurations called \emph{Pareto set} (\cite{pareto}, see also
\defR{pareto-set}). The corresponding values of the objectives for
each configuration of the Pareto Set is called \emph{Pareto front}:
it represents the trade-offs between objectives, i.e.  for each point
of the Pareto front there is no other configuration performing better
for all the objectives considered. Of course this is only a first
high-level system simulation, thus these few promising
configurations will need to be further evaluated in a next step of the design
flow using a more accurate low level simulation (e.g. register
transfer level).

\comment{DP: ho reintrodotto la descrizione di cos'e' un pareto set e
front che ci serve dopo quando descriviamo PS, inoltre ho sostituito
la tua frase era un po' contorta. In breve, il fatto che il pareto set
del DSE sia un'approssimazioen di quello divino irraggiungibile e' una
cosa che e' inutile dire qui, fa solo confondere col concetto di
approssimazione che e' invece dovuto al fatto che usiamo modelli di
alto livello}
%while providing, at the end, configurations that are close to the
%optimum and that are worth evaluating, in a following step of the
%design process, in a more thorough way.

\comment{DP: poi dobbiamo trovare un nome piu' interessante rispetto a
interesting regions, ci pensiamo questa estate}

In this work we present PS, a multiobjective design exploration
algorithm that introduces the concept of \emph{interesting} and
\emph{uninteresting regions} of the configuration space.  The main
idea can be simplified as follows: at each iteration of the algorithm,
the configuration space is divided in regions. A score is given to
each region based on the \emph{innovation} provided by its
configurations in the Pareto front. The regions with high score are
considered the most \emph{interesting} and more simulations are
distributed there rather than in the rest of the configuration space.
Several geometrical trasformations (splitting, merging etc...) are
involved in the Parameter Space (aka configuration space), hence
the name PS for the algorithm.

From a more abstract but interesting perspective, the proposed
algorithm mimics the attention shift in human beings. For example,
authors in \cite{attention} models human attention as a single
resource with a limited budget that must be distributed among tasks,
preferring the task which the focus is on.  Authors in
\cite{spatial_attention} describes the visual attention as a spotlight
that focuses, each time, in a particular region of the visual space in
order to process it more deeply. Similarly, PS focuses its
``spotlight'' only to the interesting regions and explore them more
thoroughly distributing the simulation budget on them.
% Ispirzione: http://en.wikipedia.org/wiki/Attentional_shift
In other words, with PS we propose a new perspective on the configuration space,
where parameter interdependencies play the fundamental role of making
some regions more capable of generating hardly predictable solutions:
catching this sort of ``chaoticity'' in the parameter space is the
main idea behind the strategy that is being introduced in this work.
